---
title: "Data analysis of laboratroy experiments on diploid and triploid pacific oysters"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

This is a notebook for the analysis of growth, motality, and oxygen consumption data from the single stressor experiments examining the impact of pH, temperature, and DO levels on pacific oysters. 

# Packages

```{r, results='hide'}
library(tidyverse)
library(here)
library(gridExtra)
library(lme4)
library(emmeans)
```

# Temperature

## Growth

Read in and prepare the data
```{r}
sizedat_temp <- read.csv(here("Data", "Temperature", "Size_Temperature.csv"))
sizedat_temp <- sizedat_temp %>% 
  mutate(across(c("Treatment","Tank", "Ploidy", "Time", "Tag.colour", "Tag.num"), as.factor)) 

```

Plot the size distributions of each of the ploidy groups at the start and the end of the experiment
```{r}
grid.arrange(
  sizedat_temp %>% ggplot(aes(x=Ploidy, y = Shell.Height, fill = Time))+geom_violin()+
    xlab("Ploidy")+ylab("Shell height (mm)")+theme_classic(),
  
  sizedat_temp %>% ggplot(aes(x=Ploidy, y = Mass, fill = Time))+geom_violin()+
    xlab("Ploidy")+ylab("Mass (g)")+theme_classic()
)

```

Conduct a linear mixed model to examine if there was a difference in the size of individuals at the start of the experiment and at the end. Will include each tank as a random factor. I will use the lme4 package for this.

Check the distribution of the data
```{r}
hist((sizedat_temp$Mass))
hist((sizedat_temp$Shell.Height))
```

and the distribution of log transfomed data
```{r}
hist(log(sizedat_temp$Mass))
hist(log(sizedat_temp$Shell.Height))
```

Develop the model for size and print the contrasts and p values using emmeans package
```{r}
massgrowthlme_temp <- lmer(Mass~Ploidy*Time*Treatment+ (1|Tank), data = sizedat_temp)
emmeans_massgrowthlme_temp <- emmeans(massgrowthlme_temp, ~ Ploidy+Treatment+Time)
contrast(emmeans_massgrowthlme_temp)

heightgrowthlme_temp <- lmer(Shell.Height~Ploidy*Time*Treatment + (1|Tank), data = filter(sizedat_temp))
```

## Mortality

Read in the mortality data and prepare it for analysis

```{r, results='hide'}
temp_mortdat <- read.csv(here("Data", "Temperature", "Mortality_Temperature.csv"))
str(temp_mortdat) #inspect the data
sizedat_temp <- sizedat_temp %>% #Correct the data structure
  mutate(across(c("Treatment","Tank", "Ploidy", "Time", "Tag.colour", "Tag.num"), as.factor)) 

```

# pH

## Growth

## Mortality

Read in the mortality data and prepare it for analysis. This involves changing the data from being only number of mortalities at each sampling event to giving each individual a status. This is a lot of code to do not very much

```{r}
pH_mortdat <- read.csv(here("Data", "pH", "pH_mortality.csv"))
pH_mortdattemplate <- read.csv(here("Data", "pH", "Mortality_template.csv"))
#merge the data with the tempalate
ph_mortdat_expanded <- left_join(pH_mortdattemplate, pH_mortdat, by = c("Tank", "Date", "Colour"))
#Replace the NAs with zeros
ph_mortdat_expanded <- ph_mortdat_expanded %>% mutate_all(~ifelse(is.na(.), 0, .))
#Now need to pivot wider to get the dates so we can subtract the dates
ph_mortdat_expanded <- ph_mortdat_expanded %>% pivot_wider(id_cols = c(Tank, Colour),
                                                           names_from='Date',
                                                           values_from='Number_Morts')
#Add a new collumn that has the total number per group at the start of the experiment (Think it was 15)
ph_mortdat_expanded$Survivors_Nov_2 <- 15-ph_mortdat_expanded$Nov_02

#calculate survivors at each time interval 
for (i in 4:10) {
  survivors <- ph_mortdat_expanded[, i+7] - ph_mortdat_expanded[, i]
  # Create new column name
  new_column_name <- paste0("Survivors_", colnames(ph_mortdat_expanded)[i])
  # Add new column with survivors
  ph_mortdat_expanded[, new_column_name] <- survivors
}

#now I need to create a df with 15 rows for each of the tanks and ploidys
ph_mortdat_expanded_repeated <- ph_mortdat_expanded %>%
  mutate(row_id = row_number()) %>%
  slice(rep(row_id, each = 15)) %>% 
  select(1,2)
#Now add in an id column where we repeat 1-15 for each tank
ph_mortdat_expanded_repeated$ID <- rep(1:15, length.out = nrow(ph_mortdat_expanded_repeated))
#add a status for the start of teh experiment, 0 indicates an indivvidual was alive 
ph_mortdat_expanded_repeated$status_Sep_30 <- 0

#Create a df to store all survival event data for analyes
survival_eventdat <- ph_mortdat_expanded_repeated

#create a for loop that gives all survivors a status across all
for (i in 11:18) {
  # Extract the current column name
  current_column_name <- colnames(ph_mortdat_expanded)[i]
  
  # Remove "Survivors_" from the current column name
  stripped_column_name <- gsub("Survivors_", "", current_column_name)
  
  # Create a new column name based on the stripped current column name
  new_column_name <- paste0("status_", stripped_column_name)
  
  # Uncount the specified column
  survivors_column <- ph_mortdat_expanded[c(1:2, i)] %>%
    mutate(weights = as.numeric(get(current_column_name))) %>%
    uncount(weights)
  
  # Add an ID column from 1 to 15
  survivors_column <- survivors_column %>%
    group_by(across(c(1, 2))) %>%
    mutate(ID = row_number()) %>%
    ungroup()
  
  # Add a status column to indicate that these individuals are alive
  survivors_column[[new_column_name]] <- 0
  
  # Join the data frames
  survival_eventdat <- left_join(survival_eventdat, survivors_column, by = c("Tank", "Colour", "ID"))
}
#Only keep the status columns
survival_eventdat <- survival_eventdat %>% 
  select(-starts_with("Survivors_"))
#Pivot the data longer to the correct format for survival analyses in R
survival_eventdat <- survival_eventdat %>% 
  pivot_longer(cols=c(4:12),
               names_to = "Date",
               values_to = "event") %>% 
  mutate(Date = str_replace(Date, "status_", "")) %>% #remove the word status from the name column
  mutate(Date = str_replace(Date, "Nov_", "")) %>% #replace date with experiment day
  mutate(Date = str_replace(Date, "Sep_30", "0")) %>% #replace Sep 30 with day 0
  mutate(Date = ifelse(Date == 2, "02", Date)) #replace 2 with 02

#Set the status to 1 when theree is a death (i.e. when status is NA)
survival_eventdat <- survival_eventdat %>% 
  mutate(event = ifelse(is.na(event), 1, event))

#I only want to keep the first instance of a 1 for each individual.
survival_eventdat <- survival_eventdat %>%
  group_by(ID) %>%
  mutate(event = replace(event, cumsum(event == 1) > 1, NA)) %>% 
  mutate(event = ifelse(event == 0 | cumsum(event == 1) == 1, event, 0)) %>% 
  filter(event != "NA")#Remove all NAs
#Add a treatment column to each of the tanks
#read in my excel sheet of tank assignments
pH_tank_assignments <- read.csv(here("Data", "pH", "pH_Tank_assignments.csv"))
#Add them as a treatment column to the survivor data
survival_eventdat1 <- merge(survival_eventdat, pH_tank_assignments, by = "Tank")


```

### Plot the data to visualise mortality between tanks


## Respiration

# DO 

## Growth

## Mortality

## Respiration