---
title: "Data analysis of laboratroy experiments on diploid and triploid pacific oysters"
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
---

This is a notebook for the analysis of growth, motality, and oxygen consumption data from the single stressor experiments examining the impact of pH, temperature, and DO levels on pacific oysters. 

# Packages

```{r, results='hide'}
library(tidyverse)
library(here)
library(survival)
library(ggfortify)
library(ggsurvfit)
library(survminer)
library(gridExtra)
library(frailtyEM)
library(lme4)
library(emmeans)
```

# Temperature

## Growth

Read in and prepare the data
```{r}
sizedat_temp <- read.csv(here("Data", "Temperature", "Size_Temperature.csv"))
sizedat_temp <- sizedat_temp %>% 
  mutate(across(c("Treatment","Tank", "Ploidy", "Time", "Tag.colour", "Tag.num"), as.factor)) 

```

Plot the size distributions of each of the ploidy groups at the start and the end of the experiment
```{r, echo=FALSE}
grid.arrange(
  sizedat_temp %>% ggplot(aes(x=Ploidy, y = Shell.Height, fill = Time))+geom_violin()+
    xlab("Ploidy")+ylab("Shell height (mm)")+theme_classic(),
  
  sizedat_temp %>% ggplot(aes(x=Ploidy, y = Mass, fill = Time))+geom_violin()+
    xlab("Ploidy")+ylab("Mass (g)")+theme_classic()
)

```

Conduct a linear mixed model to examine if there was a difference in the size of individuals at the start of the experiment and at the end. Will include each tank as a random factor. I will use the lme4 package for this.

Check the distribution of the data
```{r, echo=FALSE}
hist((sizedat_temp$Mass))
hist((sizedat_temp$Shell.Height))
```

and the distribution of log transfomed data. Log transformed data is more normaly distrubuted so I am going to do the analyses on log transformed data. 
```{r, echo=FALSE}
hist(log(sizedat_temp$Mass))
hist(log(sizedat_temp$Shell.Height))
```

Develop the model for size and print the contrasts and p values using emmeans package. First I am going to examine growth in terms of total oyste mass. Tank is included as a random effect to account for any inter tank variation. 
```{r}
massgrowthlme_temp <- lmer(log(Mass)~Ploidy*Time*Treatment+ (1|Tank), data = sizedat_temp) #This model includes a ransdom effect of tank. 
emmeans_massgrowthlme_temp <- emmeans(massgrowthlme_temp, ~ Ploidy+Treatment+Time)
masscontrasts <- contrast(emmeans_massgrowthlme_temp)
masscontrastsdf <-  as.data.frame(masscontrasts)
print(masscontrastsdf %>% filter(p.value <=0.05))

heightgrowthlme_temp <- lmer(Shell.Height~Ploidy*Time*Treatment + (1|Tank), data = filter(sizedat_temp))
```

The second step is to examine growth in terms of shell height. 
```{r}
heightgrowthlme_temp <- lmer(log(Shell.Height)~Ploidy*Time*Treatment + (1|Tank), data = filter(sizedat_temp))
emmeans_heightgrowthlme_temp <- emmeans(heightgrowthlme_temp, ~ Ploidy+Treatment+Time)
heightcontrasts <- contrast(emmeans_heightgrowthlme_temp)
heightcontrasts <-  as.data.frame(heightcontrasts)#Change to a df so that I can filter out the results and only show p<0.05
print(heightcontrasts %>% filter(p.value <=0.05))
```

The results indicate that significant differences existed in the size of.....

## Mortality

Read in the mortality data and prepare it for analysis

```{r, results='hide', echo=FALSE}
temp_survivors <- read.csv(here("Data", "Temperature", "Temp__survivors.csv"))
temp_survivors$Tank <- as.factor(temp_survivors$Tank)
temp_tank_assignments <- read.csv(here("Data", "Temperature", "Temp_tank_assignments.csv"))
#Uncount each of the rows in the first survivor colum
individ_mortdat_temp <- temp_survivors %>%
  select(1:3) %>%
  uncount(weights = Surv_1)
individ_mortdat_temp$ID <- rep(1:30, length.out = nrow(individ_mortdat_temp))#Add an ID column so each individual has a unique ID

individ_mortdat_temp$status_Surv_0 <- 0

#Create a df to store all survival event data for analyes
survival_eventdat_temp <- individ_mortdat_temp

#create a for loop that gives all survivors a status across all
for (i in 3:31) {
  # Extract the current column name
  current_column_name <- colnames(temp_survivors)[i]
  
  # Remove "Survivors_" from the current column name
  stripped_column_name <- current_column_name
  
  # Create a new column name based on the stripped current column name
  new_column_name <- paste0("status_", stripped_column_name)
  
  # Uncount the specified column
  survivors_column <- temp_survivors[c(1:2, i)] %>%
    mutate(weights = as.numeric(get(current_column_name))) %>%
    uncount(weights)
  
  # Add an ID column from 1 to 15
  survivors_column <- survivors_column %>%
    group_by(across(c(1, 2))) %>%
    mutate(ID = row_number()) %>%
    ungroup()
  
  # Add a status column to indicate that these individuals are alive
  survivors_column[[new_column_name]] <- 0
  
  # Join the data frames
  survival_eventdat_temp <- left_join(survival_eventdat_temp, survivors_column, by = c("Tank", "Colour", "ID"))
}
 
#Select the columns that status at the begining
survival_eventdat_temp <- survival_eventdat_temp %>% select(-starts_with("Surv"))
#Now need to pivot longer for survival plots
survival_eventdat_temp <- survival_eventdat_temp %>% pivot_longer(cols= -c(1:3), names_to = "Time", values_to = "Status")
#Remove the status from the time column
survival_eventdat_temp <- survival_eventdat_temp %>% mutate(Time = gsub("status_Surv_", "", Time))
#Add 1 where there is a NA
survival_eventdat_temp$Status[is.na(survival_eventdat_temp$Status)] <- 1
#Only keep the first instance of 1 (An individual can only die once - Cue James Bond music).
survival_eventdat_temp <- survival_eventdat_temp %>%
  group_by(Tank, Colour, ID) %>%
  filter(row_number() <= which.max(Status == 1) | sum(Status == 1) == 0) %>% ungroup()
#Make sure that the sum of mortality events actually matches the number of mortalities. In the temperature experiment this was 20
survival_eventdat_temp$event <- survival_eventdat_temp$Status
survival_eventdat_temp$Time <- as.numeric(survival_eventdat_temp$Time)
sum(survival_eventdat_temp$Status)
#Merge the file with the list of treatment tanks
survival_eventdat_temp <- merge(survival_eventdat_temp, temp_tank_assignments, by = "Tank")
```

Now the the data is in the correct format I can ployt the mortality. Beacuse mortality was so low this will not be a very interesting plot. This shows pooled mortality across temperature treatments

```{r, echo=FALSE}
pooled_site_survplot_temp_pooled <- survival_eventdat_temp %>% 
 ggsurvplot_facet(survfit(Surv(Time, event, type = "right") ~ Colour , data = .),
            facet.by = "Temp",
            data = .,
            risk.table = FALSE, pval = FALSE, conf.int = TRUE, 
            #legend.labs = c("Diploid", "Induced Triploid", "Mated Triploid"),
            xlab = "Time (days)",
             ylim = c(0.6, 1),
            legend.strata = FALSE, 
            legend.title = "Ploidy")+
  theme_bw(base_size = 15)+
  labs(legend = "ploidy")+
  scale_color_manual(values = c("blue", "green", "red"))+
  scale_fill_manual(values = c("blue", "green", "red"))
pooled_site_survplot_temp_pooled
```

Plot the survival across tanks

```{r, echo=FALSE}
pooled_site_survplot_temp_Tank <- survival_eventdat_temp %>% 
 ggsurvplot_facet(survfit(Surv(Time, event, type = "right") ~ Colour , data = .),
            facet.by = "Tank",
            data = .,
            risk.table = FALSE, pval = FALSE, conf.int = TRUE, 
            #legend.labs = c("Diploid", "Induced Triploid", "Mated Triploid"),
            xlab = "Time (days)",
             ylim = c(0.6, 1),
            legend.strata = FALSE, 
            legend.title = "Ploidy")+
  theme_bw(base_size = 15)+
  labs(legend = "ploidy")+
  scale_color_manual(values = c("blue", "green", "red"))+
  scale_fill_manual(values = c("blue", "green", "red"))
pooled_site_survplot_temp_Tank
```

And finally the same plot, except faceted by Ploidy.

```{r, echo=FALSE}
pooled_site_survplot_temp_Ploidy <- survival_eventdat_temp %>% 
 ggsurvplot_facet(survfit(Surv(Time, event, type = "right") ~ Temp , data = .),
            facet.by = "Colour",
            data = .,
            risk.table = FALSE, pval = FALSE, conf.int = TRUE, 
            #legend.labs = c("Diploid", "Induced Triploid", "Mated Triploid"),
            xlab = "Time (days)",
             #ylim = c(0.6, 1),
            legend.strata = FALSE, 
            legend.title = "Ploidy")+
  theme_bw(base_size = 15)
  #labs(legend = "ploidy")+
  #scale_color_manual(values = c("blue", "green", "red"))+
  #scale_fill_manual(values = c("blue", "green", "red"))
pooled_site_survplot_temp_Ploidy
```

## Respiration


# pH

## Growth

There was a lot of mortality across the tanks so this may be hard to work with. Regardless, I'll follow the same analysis techniques that I used for the temperature growth data. 

## Mortality

Read in the mortality data and prepare it for analysis. This involves changing the data from being only number of moralities at each sampling event to giving each individual a status. This is a lot of code to do not very much

```{r, echo=FALSE}
pH_mortdat <- read.csv(here("Data", "pH", "pH_mortality.csv"))
pH_mortdattemplate <- read.csv(here("Data", "pH", "Mortality_template.csv"))
pH_tanktrt <- read.csv(here("Data", "pH", "pH_Tank_assignments.csv"))
pH_mortdat$Tank <- as.factor(pH_mortdat$Tank)
pH_mortdattemplate$Tank <- as.factor(pH_mortdattemplate$Tank)
#merge the data with the tempalate
ph_mortdat_expanded <- left_join(pH_mortdattemplate, pH_mortdat, by = c("Tank", "Date", "Colour"))
#Replace the NAs with zeros
ph_mortdat_expanded <- ph_mortdat_expanded %>%
  mutate(Number_Morts = ifelse(is.na(Number_Morts), 0, Number_Morts))
#Now need to pivot wider to get the dates so we can subtract the dates
ph_mortdat_expanded <- ph_mortdat_expanded %>% pivot_wider(id_cols = c(Tank, Colour),
                                                           names_from='Date',
                                                           values_from='Number_Morts')
#Add a new collumn that has the total number per group at the start of the experiment (Think it was 15)
ph_mortdat_expanded$Survivors_Nov_2 <- 15-ph_mortdat_expanded$Nov_02

#calculate survivors at each time interval 
for (i in 4:10) {
  survivors <- ph_mortdat_expanded[, i+7] - ph_mortdat_expanded[, i]
  # Create new column name
  new_column_name <- paste0("Survivors_", colnames(ph_mortdat_expanded)[i])
  # Add new column with survivors
  ph_mortdat_expanded[, new_column_name] <- survivors
}

#now I need to create a df with 15 rows for each of the tanks and ploidys
ph_mortdat_expanded_repeated <- ph_mortdat_expanded %>%
  mutate(row_id = row_number()) %>%
  slice(rep(row_id, each = 15)) %>% 
  select(1,2)
#Now add in an id column where we repeat 1-15 for each tank
ph_mortdat_expanded_repeated$ID <- rep(1:15, length.out = nrow(ph_mortdat_expanded_repeated))
#add a status for the start of teh experiment, 0 indicates an indivvidual was alive 
ph_mortdat_expanded_repeated$status_Sep_30 <- 0

#Create a df to store all survival event data for analyes####################I FUCKED THIS UP SOMEHOW, GO BACK AND FIX IT#################!!!!!
for (i in 11:18) {
  # Extract the current column name
  current_column_name <- colnames(ph_mortdat_expanded)[i]
  
  # Remove "Survivors_" from the current column name
  stripped_column_name <- gsub("Survivors_", "", current_column_name)#Here I should remove "surv" from name
  
  # Create a new column name based on the stripped current column name
  new_column_name <- paste0("status_", stripped_column_name)
  
  # Uncount the specified column ###################THIS IS WHERE THINGS GO WRONG HERE###################
  survivors_column <- ph_mortdat_expanded[c(1:2, i)] %>%
    mutate(weights = as.numeric(get(current_column_name))) %>%
    uncount(weights)
  
  # Add an ID column from 1 to 15
  survivors_column <- survivors_column %>%
    group_by(across(c(1, 2))) %>%
    mutate(ID = row_number()) %>%
    ungroup()
  
  # Add a status column to indicate that these individuals are alive
  survivors_column[[new_column_name]] <- 0
  
  # Join the data frames
  ph_mortdat_expanded_repeated <- left_join(ph_mortdat_expanded_repeated, survivors_column, by = c("Tank", "Colour", "ID"))
}

#Now select only the coluimns that have the status
survival_eventdat_pH <- ph_mortdat_expanded_repeated %>% select(-starts_with("Surv"))
#Pivot longer 
survival_eventdat_pH <- survival_eventdat_pH %>% pivot_longer(cols= -c(1:3), names_to = "Time", values_to = "Status")
#Remove the added dates from the time columns
survival_eventdat_pH <- survival_eventdat_pH %>% mutate(Time = gsub("status_Nov_", "", Time))
survival_eventdat_pH <- survival_eventdat_pH %>% mutate(Time = gsub("status_Sep_30", "0", Time))
#Add 1 where there is a NA
survival_eventdat_pH$Status[is.na(survival_eventdat_pH$Status)] <- 1
#Only keep the first instance of 1 (An individual can only die once - Cue James Bond music).
survival_eventdat_pH <- survival_eventdat_pH %>%
  group_by(Tank, Colour, ID) %>%
  filter(row_number() <= which.max(Status == 1) | sum(Status == 1) == 0) %>% ungroup()
#Add an "event column
survival_eventdat_pH$event <- survival_eventdat_pH$Status
survival_eventdat_pH$Date <- as.numeric(survival_eventdat_pH$Time)
#Make sure that the sum of mortality events actually matches the number of mortalities. In the temperature experiment this was 250
sum(survival_eventdat_pH$Status)
#Join the CO2 data so we have a CO2 Level for each of our variables
survival_eventdat_pH <- merge(survival_eventdat_pH,pH_tanktrt, by = "Tank")

```

### Plot the data to visualise mortality between tanks

This plot shows pooled survival across pH treatments. This does not account for any random effects of tank that could influence survival. Just a starting point to get an overview of the data. 

```{r, echo=FALSE}
pooled_site_survplot_pH_pooled <- survival_eventdat_pH %>% 
 ggsurvplot_facet(survfit(Surv(Date, event, type = "right") ~ Colour , data = .),
            facet.by = "CO2_level",
            data = .,
            risk.table = FALSE, pval = FALSE, conf.int = TRUE, 
            #legend.labs = c("Diploid", "Induced Triploid", "Mated Triploid"),
            xlab = "Time (days)",
             ylim = c(0.6, 1),
            legend.strata = FALSE, 
            legend.title = "Ploidy")+
  theme_bw(base_size = 15)+
  labs(legend = "ploidy")+
  scale_color_manual(values = c("blue", "green", "red"))+
  scale_fill_manual(values = c("blue", "green", "red"))
pooled_site_survplot_pH_pooled
```

Here I am going to see how different ploidies reacted to the same pH treatment. The same data as above (i.e. pooled) but faceted by ploidy rather than CO2

```{r, echo=FALSE}
pooled_site_survplot_pH_ploidy <- survival_eventdat_pH %>% 
 ggsurvplot_facet(survfit(Surv(Date, event, type = "right") ~ CO2_level , data = .),
            facet.by = "Colour",
            data = .,
            risk.table = FALSE, pval = FALSE, conf.int = TRUE,
            ylim = c(0.6, 1),
            #legend.labs = c("Diploid", "Induced Triploid", "Mated Triploid"),
            xlab = "Time (days)",
            legend.strata = FALSE,
            legend.title = "pCO2")+
  theme_bw(base_size = 15)
  #labs(legend = "ploidy")+
  #scale_color_manual(values = c("blue", "green", "red"))+
  #scale_fill_manual(values = c("blue", "green", "red"))
pooled_site_survplot_pH_ploidy
```


To see how survival varied across tanks this is a plot that shows the survival probability for each ploidy group across tanks.

```{r, echo=FALSE}
pooled_site_survplot_pH_tank <- survival_eventdat_pH %>% 
 ggsurvplot_facet(survfit(Surv(Date, event, type = "right") ~ Colour , data = .),
            facet.by = "Tank",
            data = .,
            risk.table = FALSE, pval = FALSE, conf.int = TRUE, 
            #legend.labs = c("Diploid", "Induced Triploid", "Mated Triploid"),
            xlab = "Time (days)",
             ylim = c(0.6, 1),
            legend.strata = FALSE, 
            legend.title = "Ploidy")+
  theme_bw(base_size = 15)+
  labs(legend = "ploidy")+
  scale_color_manual(values = c("blue", "green", "red"))+
  scale_fill_manual(values = c("blue", "green", "red"))
pooled_site_survplot_pH_tank
```

### Statistical analysis on the mortality data

First I am going to do a pooled analysis on all the data. This does not include a random effect of tank which I will include later. Using a cox proportional hazards model based on the analysis of McElhany. 

```{r, echo=FALSE}
#Create a survival object for analyses
surv <- with(survival_eventdat_pH, Surv(Date, event, type = "right"))#For some reason piping does not work here
#Create the survival models
K1 <- coxph(surv ~ CO2_level , data = survival_eventdat_pH)
K2 <- coxph(surv ~ Colour , data = survival_eventdat_pH)
K3 <- coxph(surv ~ Colour + CO2_level, data = survival_eventdat_pH)
K4 <- coxph(surv ~ Colour * CO2_level, data = survival_eventdat_pH)

summary(K1)
summary(K2)
summary(K3)
summary(K4)

```

Results of these analyses show that there was as significant impact of pCO2 and ploidy on survival (when pooled). Significant interactions existed between ploidy and pCO2 levels. The graphs show that blue did not survive well in the highest pH treatement. 

I am going to add a random effect of tank here to assess inter tank variation. I am using the frailtyEM package (Balan and Putter 2019) as in McElhany crab survival analysis. I am going to include treatment and ploidy as fixed effects tank as a random effect. I have no idea if this is correct or how to interpret it. 

```{r}
survival_eventdat_pH1 <- survival_eventdat_pH %>% mutate(across(c("CO2_level","Tank", "Colour"), as.factor)) 
survival_eventdat_pH1$Time <- as.numeric(survival_eventdat_pH1$Time)
survival_eventdat_pH1$Time <- survival_eventdat_pH1$Time+1
pH_frailty_fit <- survival_eventdat_pH1 %>% emfrail(Surv(Time, event, type = "right") ~ CO2_level*Colour + cluster(Tank),
distribution = emfrail_dist(dist = "gamma"),
data = .)

```

## Respiration

# DO 

## Growth

## Mortality

Read in the mortality data, and format it so that I can run survival analyses. 
```{r}
DO_mortdat <- read.csv(here("Data", "DO", "DO_morts.csv"))
DO_tank_assignments <- read.csv(here("Data", "DO", "DO_Tank_assignments.csv"))



```

## Respiration